---
title: "Model specification"
author: "Marius Barth"
output: html_document
---

# Model specification {#appendix-model-specification}

We implemented the drift-diffusion model in JAGS [@plummer_jags_2003] and fit the model to each between-subjects condition, separately.
Response times and accuracy on each trial $n$ were modeled as coming from a Wiener distribution with
the models' *core* parameters
boundary separation $\alpha_{ik^*b}$, starting point $\beta_n$, drift rate $\delta_n$ and nondecision time $\tau_n$

$$y_n \sim \mathcal{W} \left(\alpha_{ik^*b}, \beta_n, \delta_n, \tau_n \right)$$

Boundary separation is assumed to vary by
participant $i$
and block pair $b$.
For mixed-deterministic materials, it also varies by condition $k^*$, denoting deterministic vs. random blocks.

$$\alpha_{ik^*b} = \zeta^{(\alpha)}_{ik^*b}$$

The other core parameter are assumed to vary by trial,
and are assumed to come from truncated normal distributions with standard deviations $s_i$.

$$\beta_n \sim \mathrm{N}_{.01}^{.99} \left( \zeta^{(\beta)}_{ikb} + p^{(\beta)} \kappa_{ij}^{(\beta)}, s^{(\beta)}_i \right)$$

$$\delta_n \sim \mathrm{N} \left( \zeta^{(\delta)}_{ikb} + p^{(\delta)} \kappa_{ij}^{(\delta)}, s^{(\delta)}_i \right)$$

$$\tau_n  \sim \mathrm{N}_{.001} \left( \zeta^{(\theta)}_{ikb} + p^{(\theta)} \kappa_{ij}^{(\theta)} + \zeta^{(\xi)}_{il} + p^{(\xi)} \kappa_{il}^{(\xi)}, s^{(\tau)}_i \right)$$

where
$\zeta_{ikb}$ is an individual's condition (nonregular, regular, deterministic) mean in block pair $b$,
$p$ is a scaling factor,
and $\kappa_{ij}$ represent an individual's stimulus-location-dependent deviations from the individual's condition mean.

Participant-level parameters $\zeta_{ikb}$ are assumed to be normally distributed, e.g.

$$\zeta_{ikb}^{(\beta)} \sim \mathrm{N} \left( \mu_{kb}^{(\beta)}, \sigma^{(\beta)} \right)$$

where 

$$\sigma^{(\beta)} \sim q^{(\beta)} t^+_{\mathit{df} = I - 1}$$

and $q$ is a scaling factor, and $t^+_{\mathit{df}}$ is a half-$t$ distribution with $I - 1$ degrees of freedom restricted to positive values.


```{r echo = FALSE}
scale_sigma <- data.frame(
  parameter = c("alpha", "beta", "nu", "theta", "xi")
  , scaling = c(.6, .1, .6, .05, .005)
)
```

  
In the SRTT, response times and accuracy vary substantially by stimulus and/or response location.
Therefore, for each participant $i$ and stimulus location $j = 1, ..., 6$, we included normally distributed parameters $\kappa_{ij}$ to capture such differences

$$\kappa_{ij}^{(\beta)} \sim \mathrm{N} \left( \mu_j^{\kappa^{(\beta)}}, \sigma_j^{\kappa^{(\beta)}} \right)~\mathrm{for}~j = 1, ..., 5$$

and

$$\kappa_{i6}^{(\beta)} = - \sum_{j = 1}^{5} \kappa_{ij}^{(\beta)}$$

The means of these parameters are modeled as being normally distributed, where

$$\mu_j^{\kappa^{(\beta)}} \sim \mathrm{N}\left(0, 1\right)~\mathrm{for}~j = 1, ..., 5$$

and

$$\mu_{6}^{\kappa^{(\beta)}} = - \sum_{j = 1}^{5} \mu_j^{\kappa^{(\beta)}}$$

Standard deviations were modeled as coming from a half-$t$ distributions

$$\sigma_j^{\kappa^{(\beta)}} \sim t^+_{\mathit{df} = I - 1}$$

<!--
Population-level parameters

- $\lambda^{(\beta)}_k$: scale
- $\gamma^{(\beta)}_k$: shape
- $\upsilon^{(\beta)}_k$ initial limit
- $\psi^{(\beta)}_k$ mean
-->

The condition means are defined as the weighted sum of one or more shifted and stretched exponential functions $f(b)$,

$$\mu_{kb} = \sum_{m = 1}^3 w_{km} f_m(b)_{\lambda_m, \gamma_m, \upsilon_m, \iota_m}$$

<!-- and t $\psi^{(\beta)}_m$ is the average difference between initial and final limit. -->

```{r results = 'asis'}
if(is.null(params$coding)) {
  cat("We use these functions in a similar fashion as model terms in linear models:
Condition means are the weighted sum of multiple temporally-changing functions (instead of regression coefficients),
and the condition weights $w_{km}$ chosen in a way that
the first function is equivalent to an intercept term,
the second function is equivalent to the effect of regular vs. nonregular trials, and
the third function is equivalent to the effect of deterministic trials vs. the average of regular and nonregular trials;
amounting to a Helmert coding scheme.")
} else {
  cat("We use these functions in a similar fashion as model terms in linear models:
Condition means are the weighted sum of multiple temporally-changing functions (instead of regression coefficients),
and the condition weights $w_{km}$ chosen in a way that conditions means are dummy coded (with nonregular trials serving as the reference group).")
}

```





The functions (depicted in Figure\ \@ref(fig:stretched-exponential)) are defined as

$$f_m(b) = \upsilon_m + (\iota_m - \upsilon_m) e^{-(\frac{b}{\lambda_m})^{-\gamma_m}}$$

where 
$\upsilon_m$ is the initial limit,
$\iota_m$ is the final limit.
$\lambda_m$ is the temporal location of the inflection point of this function, and
$\gamma_m$ is a stretching exponent (shifting the function's steepness).


The final limit $\iota_m$ may be re-written as
<!-- $f_2 = \frac{\psi \gamma B}{\lambda   \Gamma{(\gamma^{-1})}} + \upsilon$ -->
$$\iota_m = \frac{\gamma_m \psi_m B}{\lambda_m \Gamma(\gamma_m^{-1})} + \upsilon_m$$

where $B$ is the number of block pairs, $\Gamma(.)$ is the Gamma function,
and $\psi_m$ is the *average effect*, i.e., the average difference between $f_m(b)$ and the initial limit $\upsilon_m$.

We use these average effects $\psi_m$ as the target of inference to test for overall changes (first function),
the effect of nonregular vs. regular trials (second function),
and the effect of deterministic vs. random blocks (third function).
Bayes Factors ($\mathit{BF}$) are calculated using the Savage-Dickey density ratio [@wagenmakers_bayesian_2010]
from logspline densities [@kooperberg_logspline_1992].

(ref:exponential-functions) Left panel: Shifted and stretched exponential functions for different values of $\lambda$ (location) and $\gamma$ (shape), with $\upsilon = 1$ and $\iota \in \{0, 2\}$. Right panel: The average effect $\psi$ corresponds to the area under the curve (blue) between initial limit and the function's value $f(b)$.

```{r stretched-exponential, echo = FALSE, fig.env="figure", fig.width = 9, fig.height = 4, fig.cap = "(ref:exponential-functions)"}
par(mfrow = c(1, 2), las = 1)
b <- seq(1, 7, .1)

stretched_exponential <- function(x, lambda, gamma, upsilon, iota) {
  (iota - upsilon) * exp(-((x/lambda)^-gamma)) + upsilon
}

plot.new()
plot.window(xlim = c(1, 7), ylim = c(0, 2))

out <- mapply(
  FUN = function(x, lambda, gamma, upsilon, iota, col, lty) {
    y <- stretched_exponential(x, lambda, gamma, upsilon, iota)
    lines(x, y, col = col, lty = lty)
    idx <- x == floor(x)
    points(x[idx], y[idx], col = col, pch = 16)
    
  }
  , lambda  = rep(c(4, 6), times = 4, each = 1)
  , gamma   = rep(c(4, 8), times = 2, each = 2)
  , upsilon = rep(1      , times = 8, each = 1)
  , iota    = rep(c(2, 0), times = 1, each = 4)
  , col = 1:4
  , lty = rep(1:2, times = 1, each = 4)
  , MoreArgs = list(
    x = b
  )
  , SIMPLIFY = FALSE
)
axis(side = 1)
axis(side = 2)

legend(
  "topleft"
  , legend = c(
    expression(lambda==4~gamma==4)
    , expression(lambda==6~gamma==4)
    , expression(lambda==4~gamma==8)
    , expression(lambda==6~gamma==8)
  )
  , col = 1:4
  , lty = 1
  , bty = "n"
)


plot.new()
plot.window(xlim = c(1, 7), ylim = c(0, 2))

polygon(
  x = c(b, rev(b))
  , y = c(
    stretched_exponential(b, lambda = 2.5, gamma = 8, upsilon = 1, iota = 2)
    , rep(1, length(b))
  )
  , col = "skyblue1"
  , border = NA
)
lines(x = b, y = stretched_exponential(b, lambda = 2.5, gamma = 8, upsilon = 1, iota = 2), col = 1)

axis(side = 1)
axis(side = 2)
```

